---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggnc

<!-- badges: start -->
<!-- badges: end -->

The goal of `ggnc` is to allow you to easily create maps of North Carolina Counties from flat data. Flat files will contain ids and characteristics of a county, but not the longitude and latitudes that define the parimeters of the counties.  

North Carolina is the worked example in the sf package, for this reason, ggnc as an extension package deserves special attention and polishing as a possible reference point for other projects like ggfips, ggbrazil, ggswitzerland, ggunitedstates, etc. 

Using function in ggnc, you can create the ggnorthcarolina package (or peices of it as desired) in a relatively step by step fashion - and that might be a good jumping off point for modifying to ggbrasil etc.

Here is a suggested workflow, inspect each file as you go... : 

```{r, eval = F}
devtools::create()            # creates package project structure
usethis::use_data_raw()       # creates folder to create package data
ggnc::create_dataset_script() # create reference data generating code
source("data-raw/DATASET.R")  # execute code to write reference data
devtools::document()          # make reference data available in rest package
ggnc::create_geom_script()    # create geom function and test it out
devtools::load_all("R/geom.R")
ggnc::create_stamp_script()   # create stamp function
devtools::load_all("R/stamp.R")
ggnc::create_labels_script()  # create label function
devtools::load_all("R/labels.R")
devtools::load_all()          # load all functions and try together
# adapt to use case! 
```



## Installation

You can install the development version of ggnc from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("EvaMaeRey/ggnc")
```

## Contrasting the base ggplot2 approach and the ggnc approach...

This is a basic example which shows you how to solve a common problem:

```{r example}
library(tidyverse)
library(ggnc)
## basic example code
```

The base ggplot2 approach is to use an sf object as a data input.  This sf object contains a special geometry column with information on the geographic shape information.

```{r}
class(nc)
names(nc)

nc %>% 
  ggplot() + 
  geom_sf() + 
  aes(fill = SID74)  ->
classic_approach

classic_approach
```

However, sometimes analysts will only have flat files as an initial starting point.  They won't have this special geometry column at their disposal. ggnc anticipates this starting point -- and nevertheless makes it easy to create a map. 



 

```{r}
library(tidyverse)
nc_flat %>%
  ggplot() +
  aes(fips = FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)  ->
flat_file_friendly_approach

flat_file_friendly_approach
```

# more dynamic color

```{r}
last_plot() + 
  aes(fill = SID74) +
  scale_fill_viridis_c()
```

# highlighting via in-situ subsetting

```{r}
last_plot() + 
  geom_sf_countync(county = "Ashe", color = "red", lwd = .5)


last_plot() + 
  geom_sf_countync(county = c("Alleghany", "Warren"), 
                   color = "plum", lwd = .5)
                   
```

# using stamp convenience function.

```{r}
nc_flat %>% 
  select(FIPS, BIR74:NWBIR79) %>% 
  pivot_longer(-FIPS) %>% 
  group_by(name) %>%
  mutate(value_sds = (value - mean(value, na.rm = T))/sd(value, na.rm = T)) %>%
  mutate(rank = rank(value)) %>%
  ggplot() + 
  aes(fips = FIPS) +
  aes(fill = rank) + 
  geom_sf_countync() +
  facet_wrap(~name)

nc_flat %>% 
  ggplot() + 
  aes(fips = FIPS) + 
  geom_sf_countync() +
  ggnc:::stamp_sf_nc_alamance(fill = "darkred") + 
  ggnc:::stamp_sf_nc_beaufort(fill = "green")
```

# Annotation

```{r}
ggnc::nc_flat %>%
  ggplot() +
  aes(fips = FIPS, label = NAME) +
  geom_label_nc_county()

ggnc::nc_flat %>%
 ggplot() +
 aes(fips = FIPS, label = NAME) +
 geom_sf_countync() +
 geom_label_nc_county()

 ggnc::nc_flat %>%
 ggplot() +
 aes(fips = FIPS, label = SID74, fill = SID74) +
 geom_sf_countync() +
 geom_label_nc_county(color = "oldlace")
 
 ggnc::nc_flat %>%
 ggplot() +
 aes(fips = FIPS, fill = SID74,
     label = paste0(NAME, "\n", SID74)) +
 geom_sf_countync() +
 geom_label_nc_county(lineheight = .7,
 size = 2, check_overlap= T, color = "oldlace")
```


## mostly there!


```{r}
identical(flat_file_friendly_approach, classic_approach)
library(patchwork)

flat_file_friendly_approach + classic_approach

flat_file_friendly_approach / classic_approach


```



```{r}
layer_data(classic_approach) %>% head()
layer_data(flat_file_friendly_approach) %>% head()

layer_scales(classic_approach) %>% head()
layer_scales(flat_file_friendly_approach) %>% head()

layer_grob(classic_approach)[[1]]
layer_grob(flat_file_friendly_approach)[[1]]
```

## greedy aes behavior in all sf layers makes for clunkier interface

I need to look for a better example/refine the issue.

```{r}
nc_flat %>% 
  ggplot() +
  aes(x = AREA, y = PERIMETER) +
  geom_point(aes(fips = NULL)) + 
  aes(fips = fips)
```


## might be nice to keep *all* of the reference geometries

Haven't been able to do that, throwing error. The result with partial data to be displayed.

```{r}
nc_flat %>% 
  slice_sample(n = 50, replace = F) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)

nc %>% 
  mutate(SID74 = ifelse(SID74>15, NA, SID74)) %>% 
  ggplot() + 
  geom_sf() +
  aes(fill = SID74)

nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)

nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  stamp_sf_countync(fill = "darkgrey") +
  geom_sf_countync() + 
  aes(fill = SID74)

nc_flat %>% 
  mutate(SID74 = ifelse(SID74 > 15, NA, SID74)) %>% 
  ggplot() + 
  aes(fips = FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)
```





# this right join attempt throws an error. Not sure why how to address.  But also thinking maybe inner join is the right default at least.

```{r, error=T}
compute_county_nc <- function(data, scales, county = NULL){

  reference_filtered <- reference_full
  #
  if(!is.null(county)){

    county %>% tolower() -> county

    reference_filtered %>%
      dplyr::filter(.data$county_name %>%
                      tolower() %in%
                      county) ->
      reference_filtered

  }

  # to prevent overjoining
  reference_filtered %>%
    dplyr::select("fips", "geometry", "xmin",
                  "xmax", "ymin", "ymax") ->
    reference_filtered


  data %>%
    dplyr::right_join(reference_filtered, by = join_by(fips)) %>%
    dplyr::mutate(group = -1) %>%
    dplyr::select(-fips)

}


StatCountync <- ggplot2::ggproto(`_class` = "StatCountync",
                               `_inherit` = ggplot2::Stat,
                               compute_panel = compute_county_nc,
                               default_aes = ggplot2::aes(geometry =
                                                            ggplot2::after_stat(geometry)))

geom_sf_countync <- function(
                                 mapping = NULL,
                                 data = NULL,
                                 position = "identity",
                                 na.rm = FALSE,
                                 show.legend = NA,
                                 inherit.aes = TRUE, ...
                                 ) {

                                 c(ggplot2::layer_sf(
                                   stat = StatCountync,  # proto object from step 2
                                   geom = ggplot2::GeomSf,  # inherit other behavior
                                   data = data,
                                   mapping = mapping,
                                   position = position,
                                   show.legend = show.legend,
                                   inherit.aes = inherit.aes,
                                   params = rlang::list2(na.rm = na.rm, ...)),
                                   coord_sf(default = TRUE)
                                 )

                 }

nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)


```


```
Backtrace:
 1. base (local) `<fn>`(x)
 2. ggplot2:::print.ggplot(x)
 4. ggplot2:::ggplot_build.ggplot(x)
 5. layout$train_position(data, scale_x(), scale_y())
 6. ggplot2 (local) train_position(..., self = self)
 7. self$facet$train_scales(...)
 8. ggplot2 (local) train_scales(...)
 9. ggplot2:::scale_apply(layer_data, x_vars, "train", SCALE_X, x_scales)
 ```
 
 
