---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggnc

<!-- badges: start -->
<!-- badges: end -->

The goal of ggnc is to allow you to easily create maps of North Carolina Counties from flat data. Flat files will contain ids and characteristics of a county, but not the longitude and latitudes that define the parimeters of the counties.  

North Carolina is the worked example in the sf package, for this reason, ggnc as an extension package deserves special attention and polishing as a possible reference point for other projects like ggfips, ggbrazil, ggswitzerland, ggunitedstates, etc. 

## Installation

You can install the development version of ggnc from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("EvaMaeRey/ggnc")
```

## Contrasting the base ggplot2 approach and the ggnc approach...

This is a basic example which shows you how to solve a common problem:

```{r example}
library(tidyverse)
library(ggnc)
## basic example code
```

The base ggplot2 approach is to use an sf object as a data input.  This sf object contains a special geometry column with information on the geographic shape information.

```{r}
class(nc)
names(nc)

nc %>% 
  ggplot() + 
  geom_sf() ->
classic_approach

classic_approach
```

However, sometimes analysts will only have flat files as an initial starting point.  They won't have this special geometry column at their disposal. ggnc anticipates this starting point -- and nevertheless makes it easy to create a map. 



 

```{r}
library(tidyverse)
nc_flat %>%
  ggplot() +
  aes(fips = FIPS) +
  geom_sf_countync() ->
flat_file_friendly_approach

flat_file_friendly_approach

last_plot() + 
  aes(fill = SID74) +
  scale_fill_viridis_c()

last_plot() + 
  geom_sf_countync(county = "Ashe", color = "red", lwd = .5)


last_plot() + 
  geom_sf_countync(county = c("Alleghany", "Warren"), 
                   color = "plum", lwd = .5)
                   
```


```{r}
nc_flat %>% 
  select(FIPS, BIR74:NWBIR79) %>% 
  pivot_longer(-FIPS) %>% 
  group_by(name) %>%
  mutate(value_sds = (value - mean(value, na.rm = T))/sd(value, na.rm = T)) %>%
  mutate(rank = rank(value)) %>%
  ggplot() + 
  aes(fips = FIPS) +
  aes(fill = rank) + 
  geom_sf_countync() +
  facet_wrap(~name)

nc_flat %>% 
  ggplot() + 
  aes(fips = FIPS) + 
  geom_sf_countync() +
  ggnc:::stamp_sf_nc_alamance(fill = "darkred") + 
  ggnc:::stamp_sf_nc_beaufort(fill = "green")
```

## Not there yet


```{r}
identical(flat_file_friendly_approach, classic_approach)
library(patchwork)
flat_file_friendly_approach + classic_approach



```

> Graticules
Graticules are grid lines along equal longitude (meridians) or latitude (parallels) that, depending on the projection used, often plot as curved lines on a map, giving it reference in terms of longitude and latitude. The sf function st_graticule tries to create a graticule grid for arbitrary maps. As there are infinitely many projections, there are most likely many cases where it does not succeed in doing this well, and examples of these are welcomed as sf issues.

>contains a geom specially for simple feature objects, with support for graticule white lines in the background using sf::st_graticule. Support is currently good for polygons; for lines or points, your mileage may vary.

```{r}
layer_data(classic_approach) %>% head()
layer_data(flat_file_friendly_approach) %>% head()

layer_scales(classic_approach) %>% head()
layer_scales(flat_file_friendly_approach) %>% head()

layer_grob(classic_approach)[[1]]
layer_grob(flat_file_friendly_approach)[[1]]
```

## might be nice to keep all of the reference geometries

Haven't been able to do that, throwing error. The result with partial data to be displayed.

```{r}
nc_flat %>% 
  slice_sample(n = 50, replace = F) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)

nc %>% 
  mutate(SID74 = ifelse(SID74>15, NA, SID74)) %>% 
  ggplot() + 
  geom_sf() +
  aes(fill = SID74)


nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)

nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  stamp_sf_countync(fill = "darkgrey") +
  geom_sf_countync() + 
  aes(fill = SID74)


nc_flat %>% 
  mutate(SID74 = ifelse(SID74 > 15, NA, SID74)) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)
```

# this right join attempt throws an error. Not sure how to address

```{r, error=T}
compute_county_nc <- function(data, scales, county = NULL){

  reference_filtered <- reference_full
  #
  if(!is.null(county)){

    county %>% tolower() -> county

    reference_filtered %>%
      dplyr::filter(.data$county_name %>%
                      tolower() %in%
                      county) ->
      reference_filtered

  }

  # to prevent overjoining
  reference_filtered %>%
    dplyr::select("fips", "geometry", "xmin",
                  "xmax", "ymin", "ymax") ->
    reference_filtered


  data %>%
    dplyr::right_join(reference_filtered, by = join_by(fips)) %>%
    dplyr::mutate(group = -1) %>%
    dplyr::select(-fips)

}


StatCountync <- ggplot2::ggproto(`_class` = "StatCountync",
                               `_inherit` = ggplot2::Stat,
                               compute_panel = compute_county_nc,
                               default_aes = ggplot2::aes(geometry =
                                                            ggplot2::after_stat(geometry)))

geom_sf_countync <- function(
                                 mapping = NULL,
                                 data = NULL,
                                 position = "identity",
                                 na.rm = FALSE,
                                 show.legend = NA,
                                 inherit.aes = TRUE, ...
                                 ) {

                                 c(ggplot2::layer_sf(
                                   stat = StatCountync,  # proto object from step 2
                                   geom = ggplot2::GeomSf,  # inherit other behavior
                                   data = data,
                                   mapping = mapping,
                                   position = position,
                                   show.legend = show.legend,
                                   inherit.aes = inherit.aes,
                                   params = rlang::list2(na.rm = na.rm, ...)),
                                   coord_sf(default = TRUE)
                                 )

                 }

nc_flat %>% 
  filter(SID74 <= 15) %>% 
  ggplot() + 
  aes(fips =FIPS) +
  geom_sf_countync() + 
  aes(fill = SID74)


```


```
Backtrace:
 1. base (local) `<fn>`(x)
 2. ggplot2:::print.ggplot(x)
 4. ggplot2:::ggplot_build.ggplot(x)
 5. layout$train_position(data, scale_x(), scale_y())
 6. ggplot2 (local) train_position(..., self = self)
 7. self$facet$train_scales(...)
 8. ggplot2 (local) train_scales(...)
 9. ggplot2:::scale_apply(layer_data, x_vars, "train", SCALE_X, x_scales)
 ```
